Improvement of Korean Morphological Analysis System \\Through Transformer-Based Re-Ranking

### 2023-12-02 contents/draft.tex
Apply revision by soomgo: Section 2 and 3

    In this study, we used three major corpora to train and evaluate Korean morphological analysis models, each of which has unique characteristics and serves different research purposes.

    \textbf{Sejong Corpus}: Originating from the 21st Century Sejong Project, this corpus consists of a total of 15 million eojeols, including the raw untagged corpus, and forms the backbone of Korean morphological analysis research~\cite{ChoeMW2008}.
    It provides a wide variety of linguistic patterns and structures that are important for baseline training and validation of morphological analysis models, and has been used for performance comparisons with other studies.
    In most of the previous studies, only a part of the Sejong corpus was used for experiments, and in this study, we used the dataset provided by the researchers in ~\cite{NaSH2014}.

    \textbf{UCorpus (University of Ulsan Corpus)}~\cite{UCorpusHG}: This is an extension of the Sejong corpus, which is constantly being maintained and added to by the University of Ulsan, significantly increasing its volume to 63 million eojeols and testing the adaptability and accuracy of the model to a wider range of data.
    The expansion includes corrections to previously raised errors ~\cite{KimIH2010} and a number of annotation outputs for new data, providing a substantial basis for comprehensive linguistic analysis.

    \textbf{Everyone's Corpus}~\cite{EveryoneCorpus}: Launched by the National Institute of the Korean Language in 2020, the Everybody's Corpus enriches the data landscape with web texts and spoken language materials that reflect contemporary language use~\cite{KimIH2019}.
    This modern corpus reflects the dynamic evolution of the Korean language and is playing a pivotal role in improving models for capturing the nuances of current Korean usage.

    Table~\ref{tab:data-statistics} details the specific number of sentences and words in each corpus, and the data subsets extracted for model training and evaluation.
    During the training data conversion process, we first removed duplicate sentences and excluded sentences with annotation errors or other problems.
    We can see that many duplicate sentences occur, especially in spoken language.

@@

    To train a dictionary-based morpheme analysis model effectively, the morpheme-tagged corpus, typically represented in lemma form, must be transformed to include the boundary information between morphemes in its surface form.
    Crucial to this transformation is string alignment, a process that accounts for the discrepancies between lemma forms and surface forms in the Korean morphological analysis corpus.

    In this study, string alignment was performed using the Smith-Waterman algorithm, which uses a scoring matrix based on the similarity of the grapheme unit of Korean letters for each word pair (as shown in Figure~\ref{fig:sample}). Each aligned sentence containing a morpheme tag was converted into a training sample tailored for dictionary-based morphological analysis.

    The resulting table in Figure~\ref{fig:sample} illustrates this process.
    Each row acts as a lexical unit.
    The first four columns contribute to feature generation, whereas the last four columns facilitate post-lemmatization.
    Using the morphological corpus above, a large number of training samples can be generated according to the process shown in Figure~\ref{fig:sample}.
    With the exception of the evaluation samples, the remaining sentences were used to train the dictionary-based morphological analysis model using the CRF algorithm.
    The output of this training allowed the calculation of the costs associated with each morpheme node and the linking of two consecutive morphemes.
    This, in turn, allowed the discovery of an optimal path using the Viterbi algorithm.

@@

    Figure~\ref{fig:lattice} presents a snapshot of the lattice structure, which is an integral part of the morphological analysis.
    (1) shows a fragment of the lattice structure formed when the example sentence in Figure~\ref{fig:sample} is entered.
    (2) shows the optimal path determined using the Viterbi algorithm.

    However, the path inferred by the trained model may differ from the correct solution constructed by humans.
    The nodes marked with stars in (1) represent correct nodes.
    The upper-left number of each node indicates the ranking of the nodes accessible at each decoding point.
    The choices made at certain moments deviate from the correct solution.
    To improve analytical performance, mechanisms to correct these discrepancies must developed.

@@

    Although dictionary-based morphological analysis offers significant advances, its optimal paths occasionally diverge from the correct solutions that humans understand.
    This divergence underscores the need for a model that reevaluates these primary results and reorients them to achieve higher accuracy.
    This method, called re-ranking, involves generating multiple analyses of an input and then reordering them based on a new set of criteria or models, thereby improving the overall quality of the results.

@@

    Before reranking begins, multiple analyses, typically referred to as the N-best paths, of the input sentence are generated.
    This involves extracting the top N atoms from the lattice structure.
    In this study, a novel approach is introduced to generate secondary paths, as shown in (3) of Figure~\ref{fig:lattice}, by selecting the second-best node rather than each best node constituting the path from the best-path result.
    Some of these secondary paths provided alternatives that reconciled incorrect with correct answers.
    Similarly, paths modified by favoring the third-best node were called tertiary paths, and this naming convention was continued for subsequent paths.
    In our preliminary test, the secondary paths, including the optimal and suboptimal paths, were shown to cover the majority of the correct morphological analyses as measured through human evaluations.
    (Refer to Table~\ref{tab:maximum-performance}).

@@

    Bidirectional Encoder Representations from Transformers (BERT) models~\cite{Devlin2019} have revolutionized many natural language processing tasks by understanding the context in which words appear in text.
    In this study, we attempt to leverage the power of BERT to reorder the generated secondary paths.
    We labeled the generated secondary paths with scores related to the morphological analysis performance to fine-tune a pre-trained BERT model specialized for Koreans with excess amount Korean text.
    After pre-testing several scoring methods on a modest scale, we found that using scores based on the degree of error rather than scores based on accuracy by widen the gap between correct and incorrect answers is effective for learning.

    Once the BERT model is fine-tuned and trained for the re-ranking task, it can predict a re-ranking score for each path in the secondary path list.
    This implies that considering the context, morphological organization, and other essential linguistic features of the path, the model assigns a score to each path.
    The paths were then re-ranked according to this score, and the path with the highest score was selected for the best morphological analysis.

@@

    In the first step, we re-rank the secondary paths generated using the BERT model, as described in Section ~\ref{subsec:bert-based-reranking}.
    In the second step, we introduced another BERT variant that was optimized for a different set of linguistic features or trained on a different dataset.
    This allowed us to perform a fine-grained re-evaluation, further refine the list, and push more contextually accurate paths to the top.

    As shown in Figure~\ref{fig:ranking}, for a two-stage reranking model, the first stage performs the first re-ranking, taking as input a secondary path in morphologically tagged lemma form.
    It then performs a second re-ranking, again taking as input the path re-ranked in stage 1 and the original input sentence.
    This was conducted to improve effectiveness, as it was ineffective given the same type of input.

    In summary, the re-ranking model represents a significant advancement in our approach to Korean morphological analysis.
    By leveraging BERT-based models, we anticipate a marked improvement in accuracy, especially in complex linguistic scenarios.
    The next section will present our experimental setup and results, providing crucial empirical evidence for the efficacy of the re-ranking model in practical applications.

>>>>

    In this study, three major corpora were utilized to train and evaluate Korean morphological analysis models, each serving distinct research purposes and possessing unique characteristics:

    \textbf{Sejong Corpus}: Originating from the 21st Century Sejong Project, this corpus comprises a total of 15 million eojeols, including the raw untagged corpus~\cite{ChoeMW2008}.
    It forms the backbone of Korean morphological analysis research, offering a diverse array of linguistic patterns and structures crucial for baseline training and validation of morphological analysis models.
    The Sejong Corpus has been widely used for performance comparisons with other studies.
    For our experiments, we utilized the dataset used by researchers of ~\cite{MinJW2019, MinJW2020, MinJW2022, MinJW2018, NaSH2015, NaSH2014, NaSH2018, SongHJ2019, SongHJ2020}.

    \textbf{UCorpus (University of Ulsan Corpus)}~\cite{UCorpusHG}: An extension of the Sejong corpus, the UCorpus is continually maintained and expanded by the University of Ulsan.
    It has significantly grown in volume, reaching 63 million eojeols.
    This extension tests the adaptability and accuracy of the model across a broader range of data.
    Corrections to previously identified errors ~\cite{KimIH2010} and additional annotations for new data contribute to its value, providing a comprehensive basis for linguistic analysis.

    \textbf{Everyone's Corpus}~\cite{EveryoneCorpus}: Launched by the National Institute of the Korean Language in 2020, the Everyone's Corpus enriches the data landscape with contemporary web texts and spoken language materials~\cite{KimIH2019}.
    This modern corpus reflects the dynamic evolution of the Korean language, playing a pivotal role in improving models to capture the nuances of current Korean usage.

    Table~\ref{tab:data-statistics} presents specific details regarding the number of sentences and words in each corpus, along with the data subsets used for model training and evaluation.
    In the process of converting training data, we initially removed duplicate sentences and excluded those with annotation errors or other issues.
    Notably, a substantial occurrence of duplicate sentences was observed, particularly in spoken language datasets.

@@

    To effectively train a dictionary-based morpheme analysis model, the morpheme-tagged corpus, typically represented in lemma form, needs transformation to include boundary information between morphemes in its surface form.
    This transformation relies on string alignment, addressing discrepancies between lemma forms and surface forms in the Korean morphological analysis corpus.

    In this study, we employed the Smith-Waterman algorithm for string alignment.
    This algorithm utilizes a scoring matrix based on the similarity of the grapheme unit of Korean letters for each word pair (as depicted in Figure~\ref{fig:sample}).
    Each aligned sentence containing a morpheme tag was then converted into a training sample tailored for dictionary-based morphological analysis.

    The resulting table in Figure~\ref{fig:sample} illustrates this process.
    Each row functions as a lexical unit, with the first four columns contributing to feature generation and the last four columns facilitating post-lemmatization.
    Leveraging the morphological corpus, a substantial number of training samples were generated following the process illustrated in Figure~\ref{fig:sample}.
    Except for the evaluation samples, the remaining sentences were employed to train the dictionary-based morphological analysis model using the CRF algorithm.
    The output of this training facilitated the calculation of costs associated with each morpheme node and the linking of two consecutive morphemes, enabling the determination of an optimal path using the Viterbi algorithm.

@@

    In Figure~\ref{fig:lattice}, a snapshot of the lattice structure crucial to morphological analysis is presented.
    (1) displays a portion of the lattice structure formed when inputting the example sentence from Figure~\ref{fig:sample}.
    (2) illustrates the optimal path determined through the Viterbi algorithm.

    However, it's essential to note that the path predicted by the trained model might differ from the correct solution crafted by humans.
    The nodes marked with stars in (1) represent correct nodes.
    The upper-left number of each node indicates the ranking of accessible nodes at each decoding point.
    Choices made at certain moments deviate from the correct solution.
    To enhance analytical performance, mechanisms must be developed to correct these discrepancies.

@@

    While dictionary-based morphological analysis provides substantial advancements, it is not immune to instances where its optimal paths deviate from the correct solutions perceived by humans.
    This deviation emphasizes the necessity for a model that reexamines these initial results and adjusts them to enhance accuracy.
    This approach, known as re-ranking, entails producing multiple analyses of an input and subsequently rearranging them using a new set of criteria or models, thereby elevating the overall quality of the results.

@@

    Before the re-ranking process initiates, multiple analyses, commonly referred to as N-best paths, of the input sentence are generated.
    This involves extracting the top N candidates from the lattice structure.
    In our study, a novel approach is introduced to produce secondary paths, as depicted in (3) of Figure~\ref{fig:lattice}, by selecting the second-best node instead of each best node constituting the path from the best-path result.
    Some of these secondary paths offered alternatives that reconciled incorrect answers with correct ones.
    Similarly, paths modified by favoring the third-best node were termed tertiary paths, and this nomenclature continued for subsequent paths.
    In our preliminary test, the secondary paths, encompassing both optimal and suboptimal paths, demonstrated coverage of the majority of correct morphological analyses, as assessed through human evaluations (refer to Table~\ref{tab:maximum-performance}).

@@

    Bidirectional Encoder Representations from Transformers (BERT) models have transformed numerous natural language processing tasks by comprehending the contextual nuances in which words appear in text.
    In our study, we aim to harness the capabilities of BERT to reorder the generated secondary paths.
    We assigned scores related to morphological analysis performance to the generated secondary paths and utilized them for fine-tuning a pre-trained BERT model specifically designed for Korean, enriched with a substantial amount of Korean text.
    After preliminary testing with various scoring methods on a modest scale, we found that using scores based on the degree of error, rather than accuracy-based scores, effectively widens the gap between correct and incorrect answers.

    Once the BERT model is fine-tuned and trained for the re-ranking task, it can predict a re-ranking score for each path in the secondary path list.
    This means that, taking into account the context, morphological organization, and other crucial linguistic features of the path, the model assigns a score to each path.
    Subsequently, the paths are re-ranked based on these scores, and the path with the highest score is selected as the optimal morphological analysis.

@@

    In the first step, we re-rank the secondary paths generated using the BERT model, as outlined in Section~\ref{subsec:bert-based-reranking}.
    Subsequently, in the second step, we introduce another BERT variant optimized for a different set of linguistic features or trained on a distinct dataset.
    This enables a fine-grained re-evaluation, further refining the list and elevating more contextually accurate paths to the top.

    As shown in Figure~\ref{fig:ranking}, for a two-stage re-ranking model, the first stage conducts the initial re-ranking, taking a secondary path in morphologically tagged lemma form as input.
    The second re-ranking is then performed, taking the path re-ranked in stage 1 and the original input sentence as input.
    This approach enhances effectiveness, considering the varied input types.

    In summary, the re-ranking model represents a significant advancement in our approach to Korean morphological analysis.
    By harnessing BERT-based models, we anticipate a notable improvement in accuracy, particularly in complex linguistic scenarios.
    The following section will detail our experimental setup and results, providing crucial empirical evidence for the effectiveness of the re-ranking model in practical applications.
